Scalar Parametrization for Joint Time-Frequency interpolation in a MIMO-OFDM channel

### 2019-10-09 paper/main.tex
Add back lost text
Add more changes

\section{System Model}

\label{section2}
\begin{equation}
\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
\end{equation}
where the channel matrix $H_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input $x_{i,t} \in \mathbb{C}^{N_T}$; the output $y_{i,t} \in \mathbb{C}^{N_R}$ and $\eta$ is the complex additive white gaussian noise vector with distribution $\tilde{\mathcal{N}}(0,I_{N_R})$. On applying thin SVD to H, we get $H_{i,t} = U_{i,t} \Sigma_{i,t} V_{i,t}^{H}$, where $U_{i,t} \in \mathbb{C}^{N_R \times N_T}$, $\Sigma_{i,t} \in \mathbb{C}^{N_T \times N_T}$  and $V_{i,t} \in \mathbb{C}^{N_T \times N_T}$. $U_{i,t}$ and $V_{i,t}$ are orthonormal column matrices and $\Sigma_{i,t}$ contains the singular values $\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of H. The entries of H are assumed to be i.i.d. $\tilde{\mathcal{N}}(0,1)$, complex circularly symmetric Gaussian with mean zero and variance 1. Therefore, the first n column vectors of $V_{i,t}$ have to be quantized and fed back to the transmitter as channel spatial information.

@@

\subsection{Scalar Paramatrization}
\label{givens}
The degrees of freedom in the spatial information matrix is much smaller than the number of real entries in the matrix because of the geometrical structure between the columns of the matrix~\cite{4114278}. An orthonormal matrix with $t$ rows and $n$ columns can be represented as

\begin{equation}
\textbf{V} = \Bigg[\prod_{k=1}^{n} D_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} G_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{I}
\end{equation}

where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k}, \ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( 1_{k-1}, e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, $1_{k-1}$ is (k-1) ones; and t x n matrix $\tilde{I}$ = $\big[I_{n}, 0_{n,t-n}\big]^{T}$. The Givens matrix \textbf{G} is given as

\begin{equation}
G_{m-1,m}\big(\theta\big)  =
\begin{bmatrix}

I_{m-2} & & & \\
& \cos\theta & -\sin\theta & \\
& \sin\theta & \cos\theta & \\
& & & I_{t-m}

\end{bmatrix}
\end{equation}

\begin{figure}
\label{adpm-fig}
\includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
\caption{Subcarriers which need to be quantized}
\label{ber_overvie}
\vspace{-5pt}
\end{figure}

For a $4 \times 2$ orthogonal matrix V,

$$V = D_{1}(\phi_{1,1},\ldots,\phi_{1,4}).G_{3,4}(\theta_{1,1}) .G_{2,3}(\theta_{1,2}) .G_{1,2}(\theta_{1,3})$$

\vspace{-1.4em}

\hspace{1pt}$$.D_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) .G_{3,4}(\theta_{2,1}) .G_{2,3}(\theta_{2,2}).\tilde{I}$$

Here, we denote $\phi_{i}'s$ by phases and $\theta_{i}'s$ by rotation angles. $\phi \in (-\pi, + \pi]$ is uniformaly distributed between the two extremes\cite{4114278}, while $\theta$ has a distribution

%TODO What is t in the equation below?
\begin{equation}
p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
\end{equation}
Here all $\theta_i$ and $\phi_i$ are statistically independent, making it useful for tracking them across time as explained in the next subsection. The total number of parameters obtained from the decomposition of a complex orthogonal matrix $N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
% may cross pi and come immediately to -pi, therefore, values are unwrapped while calculating them. Unwrapping is performed at the receiver over the frequency initially and later in time.

% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.

\subsection{Differential Quantization - Channel Tracking}
\label{quantiz}

For a slowly time varying channel, ADPM is an effective method for tracking a paramater over time with using only 1-bit. The 1-bit number $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$ where $x$ is the unquantized vector $\&$ $\hat{x}$ is the quantized vector and $Q(x)$ is the quantization function given as

\begin{equation}
  Q(x)=\begin{cases}
    1, & \text{if $x>0$}.\\
    -1, & \text{otherwise}.
  \end{cases}
\end{equation}

Thus, the quantied vector $\hat{x}_n$ changes with time as

\begin{equation}
\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n}
\end{equation}

\begin{equation}
\label{delta_eqn}
\Delta_{n} = \begin{cases}
    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}.\\
    \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
  \end{cases}
\end{equation}

Here we initialize delta using $\Delta_1 = abs(x_{2}-\hat{x}_1)$.

We need to quantize the subcarriers in a way such that we can exploit
the joint time-frequency interpolation at the transmitter
side. Therefore, we select the subcarriers for quantization in an
alternating fashion as shown in the figure. Figure shows a \textbf{N}
subcarrier MIMO channel with subcarriers which are going to be
quantized at the position sk, k = 0,1,...,$\frac{N-1}{s}$ for even
time instance. Here, s is the gap between two frequencies. For odd
time instances, subcarriers at the position sk + t for k = 0,1,...,
$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$.

@@

\bf{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} . This ratio was decided after trying with different ratios. This implies that if quantied values loose track of the actual values then after changing delta quickly they start following it again.
Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare

>>>>

\section{System Model}
For a MIMO-OFDM system with $N_T$ transmitter and $N_R$ receiver
antennas, each subcarrier $i$ with time instant $t$ could be modelled as a flat-fading MIMO channel. This flat-fading MIMO channel can be modelled as
\label{section2}
\begin{equation}
\by_{i,t} = \bH_{i,t} \bx_{i,t}+ \bf{\eta_{i,t}}
\end{equation}
where the channel matrix is
$\bH_{i,t} \in \mathbb{C}^{N_R \times N_T}$; the input is
$\bx_{i,t} \in \mathbb{C}^{N_T}$; the output is
$\by_{i,t} \in \mathbb{C}^{N_R}$ and the complex additive white
Gaussian noise is $\eta$ whose distribution is
${\mathcal{CN}}(0,I_{N_R})$. We consider the case where $N_R <
N_T$. The thin SVD of $\bH$ is given by
$\bH_{i,t} = \bU_{i,t} \Sigma_{i,t} \bV_{i,t}^{H}$, where
$\bU_{i,t} \in \mathbb{C}^{N_R \times N_R}$,
$\Sigma_{i,t} \in \mathbb{C}^{N_R \times N_R}$ and
$\bV_{i,t} \in \mathbb{C}^{N_T \times N_R}$. $\bU_{i,t}$ and
$\bV_{i,t}$ are matrices whose columns are orthonormal, and
$\Sigma_{i,t}$ contains the singular values
$\sigma_1 \geq \ldots \geq \sigma_{N_T} > 0$ of $\bH_{i,t}$. The
entries of $\bH_{i,t}$ are assumed to be distributed
i.i.d. $\tilde{\mathcal{N}}(0,1)$. To obtain a precoder at the
transmitter, we need to quantize and feed back information about
$\bV_{i,t}$ for each subcarrier.

@@

\subsection{Scalar Parametrization}
\label{givens}
The degrees of freedom in the $\bV_{i,t}$ matrices is smaller than the
number of real entries in the matrix because of the geometrical
structure between the columns of the matrix~\cite{4114278}. The
orthonormal matrix $\bV_{i,t}$ with $N_T$ rows and $N_R$ columns can be
decomposed as TODO
\begin{equation}
\textbf{V} = \Bigg[\prod_{k=1}^{n} \bD_{k} \big( \phi_{k,k},\ldots , \phi_{k,n} \big) \: \Bigg\{ \prod_{l=1}^{l=t-k} \bG_{t-l,t-l+1} \big( \theta_{k,l}\big) \Bigg\} \Bigg] \: \tilde{\bI}
\end{equation}
where $D_{k}$ is the diagonal matrix given as $D_{k}\big(\phi_{k,k},
\ldots, \phi_{k,t} \big)$ =  $\mbox{diag}\big( {\bf 1}_{k-1},
e^{j\phi_{k,k}},\ldots, e^{j\phi_{k,t}}  \big)$, ${\bf 1}_{k-1}$ is
represents  $k-1$ ones, and the $N_R \times N_T$ matrix $\tilde{\bI}$
= $\big[\bI_{n}, {\bf 0}_{n,t-n}\big]^{T}$, and
\begin{equation}
G_{m-1,m}\big(\theta\big)  =
\begin{bmatrix}
I_{m-2} & & & \\
& \cos\theta & -\sin\theta & \\
& \sin\theta & \cos\theta & \\
& & & I_{t-m}
\end{bmatrix}
\end{equation}
Thus, for a $4 \times 2$ orthogonal matrix $\bV_{i,t}$, we have
\begin{align*}
  \bV_{i,t} & =
  \bD_{1}(\phi_{1,1},\ldots,\phi_{1,4})\bG_{3,4}(\theta_{1,1})
  \bG_{2,3}(\theta_{1,2}) \bG_{1,2}(\theta_{1,3})\times\\
& \bD_{2}(\phi_{2,2},\phi_{2,3},\phi_{2,4}) \bG_{3,4}(\theta_{2,1}) \bG_{2,3}(\theta_{2,2})\tilde{\bI}
\end{align*}
Here, $\phi_{k}$s are referred to as phases and $\theta_{l}$s as
rotation angles. $\phi_k \in (-\pi, + \pi]$ is uniformly distributed
between the two extremes\cite{4114278}, while $\theta_l$
is distributed as
\begin{equation}
p(\theta) = 2l(\sin\theta)^{2l-1}\cos\theta, \theta \in \left[0, \frac{\pi}{2}\right) \mbox{ for } l = 1,2,\ldots,t-1
\end{equation}
The $\phi_k$ and $\theta_l$ are statistically independent, making this
representation useful for tracking them separately across time, as
explained in the following subsection. The total number of parameters
obtained from the decomposition of a complex orthogonal matrix
$N_{T} \times N_{R} $ is $N_{R}(2N_{T} - N_{R})$~\cite{4114278}. The
number of $\phi_i$ parameters is $N_{R}(2N_{T} - N_{R}-1)/2$ while the
number of $\theta$ parameters is $N_{T}(2N_{T} - N_{R}+1)/2$.
\begin{figure}
\includegraphics[width=0.45\textwidth]{images/new-adpm.pdf}
\caption{\label{fig:adpm-fig}Subcarriers which need to be quantized}
\label{ber_overvie}
\vspace{-5pt}
\end{figure}
% To exploit the interpolation of time \& frequency jointly at the transmitter, feedback bits are sent in an alternating fashion as shown in the figure.

\subsection{Differential Quantization - Channel Tracking}
\label{quantiz}
For a channel that varies slowly with time, ADPM is an effective
method for tracking parameters over time using very few bits (often
just one bit). For example, while tracking a process $x_n$, at time
$n$, the one-bit jump $\beta_{n} = Q(x_{n} - \hat{x}_{n-1})$, where
$x_n$ is the unquantized value and $\hat{x}_{n-1}$ is the unquantized
(accurate) value, the quantization function $Q(x)$ is given as
\begin{equation}
  Q(x)=\begin{cases}
    1, & \text{if $x>0$}.\\
    -1, & \text{otherwise}.
  \end{cases}
\end{equation}
Thus, the quantied vector $\hat{x}_n$ evolves with time as
\begin{equation}
\hat{x}_{n} = \hat{x}_{n-1} + \beta_{n}\Delta_{n} \mbox{ where }
\label{delta_eqn}
\Delta_{n} = \begin{cases}
    M \Delta_{n-1}, & \text{if $\beta_{n} = \beta_{n-1}$}\\
    \Delta_{n-1}/m , & \text{if $\beta_{n} \neq \beta_{n-1}$}.
  \end{cases}
\end{equation}
Here we initialize delta using $\Delta_1 = |x_{2}-\hat{x}_1|$.

To quantize the subcarriers efficiently and exploit the joint
time-frequency interpolation at the transmitter side, we select the
subcarriers for quantization in an alternating fashion as shown in
Fig.~\ref{fig:adpm-fig}. The figure shows the channel evolution of an
$N$ subcarrier MIMO-OFDM system with subcarriers which are going to be
quantized at every $s$-th subcarriers, for even time instances, where,
$s$ is the gap between two quantized subcarrier indices. For odd time
instances, subcarriers at the position $sk$ for k = 0,1,...,
$\frac{N-1}{s}$ with t = ${\frac{s}{2}}$. TODO

@@

{ADPM Design}: The decrement rate - \bm is 2.4 times the \bM - increment rate in the Equation~\ref{delta_eqn} . This ratio was decided after trying with different ratios. This implies that if quantied values lose track of the actual values then after changing delta quickly they start following it again.
Also, need to describe the initialization of the Delta values required in the quantization process. Need to cite that paper. We should also try over different quantization rates to compare
